# Engineering Statistics
IMSE 4410 Spring 2015. 
Copyright 2013-2015 by Timothy Middelkoop License CC by SA 3.0

## Goodness of Fit (Chapter 9-7, Week 8, Graduate)

* What if the underlying distribution is unknown?
* Collect $n$ samples and place them in $k$ bins or intervals (can be unequal).
* $O_i$ is the observed frequency.
* $E_i$ is the expected frequency of the hypothesized distribution with $p$ parameters.
* The test statistic follows the $X^2$ (Chi-sq) distribution and is
* $X^2_0=\sum_{i=0}^k {(O_i-E_i)^2 \over E_i}$
* with $k-p-1$ degrees of freedom.
* There are issues when the frequencies are small (see text).

```{r}
# Example 9-12: Circuit Board Defects

# Hypothesized distribution: Poisson distribution
# Sample:
n <- 60
d <- data.frame(defects=c(0,1,2,3),frequency=c(32,15,9,4))
d

# The sum of the frequencies and n should match.
sum(d$frequency) ## [1] 60

# Make the row names match to reduce the possibility of error.
rownames(d) <- d$defects ; d
```

We have $O_i$, we now need to build $E_i$

We estimate the mean of the Poisson distribution from the sample mean (xb)
$$\bar x = E(X)=\lambda T$$ where T is the unit basis, in this case Single Board

```{r}
xb <- sum(d$defects*d$frequency)/n ; xb
# [1] 0.75

# We use this to compute the expected probability, 
# and by extension the frequency, we will do this directly using R.

# Compute the bin frequencies.
d$p <- dpois(d$defects,lambda=xb) ; d$p

# We note the last one should contain all frequencies, P(X>=3)
# To keep things straight we use the label, not the row number 
d['3',]$p <- 1-sum(dpois(0:2,lambda=xb))
d['3',]

# We can also get to it by value
d[d$defects==3,]

# check our answer
sum(d$p) ## [1] 1

# now compute Ei
d$expected <- d$p*n ; d 

# Following the text we need to combine 2 and 3 since bin 3 is small.
# Create a new table with correct number of rows
h <- d[d$defects <= 2,]

# Last two rows
d[d$defects >= 2,]

sum(d[d$defects >= 2,]) ## [1] 28.57487

# Combine last two rows, replacing last row.
h['2',] <- colSums(d[d$defects >= 2,]) ; h

# Fixup the defects column
h['2',]$defects <- 2 ; h

# This matches the text.

# Check that we did everything fine.
colSums(h)
```

Note if this was a repeated analysis you could easily make functions to do all this.

Now for the hypothesis test.

* H0: it is a Poisson distribution. 
* H1: it is not a Poisson distribution.
* Test Statistic: $X^2_0=\sum (oi-Ei)^2/Ei$
* Reject H0 if p-value <0.05 (alpha), 3-1-1 degrees of freedom, upper tail
* Calculations:

```{r}
chisq0 <- sum((h$frequency-h$expected)^2/h$expected) ; chisq0 ## [1] 2.962772

pv <- pchisq(chisq0,nrow(h)-1-1,lower.tail=FALSE) ; pv ## [1] 0.08520178
```

pv > 0.05 so we fail to reject H0

There is no strong evidence to say the distribution is not Poisson.

## Example 9-13.

* Parameter of interest: form of the distribution.
* H0: the distribution is normal
* H1: the distribution is not normal
* Test statistic: chisq0=sum((oi-Ei)^2/Ei
* Reject H0 if: pv < 0.05 (alpha), upper tail, with k-p-1 dof.
* Calculate:

```{r}
d <- data.frame(observed=c(12,14,12,13,12,11,12,14),expected=c(12.5)) ; d
chisq0 <- sum((d$observed-d$expected)^2/d$expected) ; chisq0 ## [1] 0.64

# degrees of freedom (k-p-1)
dof <- nrow(d)-2-1 ; dof ## [1] 5

pv <- pchisq(chisq0,dof,lower.tail=FALSE) ; pv ## [1] 0.9860987
```

* Result: pv > 0.05 so we fail to reject H0
* Conclusions: There is no strong evidence to say the distribution is not normal

```{r}
# How do we construct the  interval class
xb <- 5.04 ; s <- 0.08; n <- 100
k <- 8 ; ei <- n*1/k ; ei

# Compute the interval (check with text)
qnorm(1/k,mean=xb,sd=s,lower.tail=TRUE) ## [1] 4.947972

# We now need a sequence to generate our intervals
#?seq
seq(0,1,by=1/k)

# generate our intervals and data.frame.
d <- data.frame(pl=seq(0,1-1/k,by=1/k),pu=seq(0+1/k,1,by=1/k)) ; d
d$lower <- qnorm(d$pl,mean=xb,sd=s)
d$upper <- qnorm(d$pu,mean=xb,sd=s) 
d
```

