---
output: pdf_document
---
# Engineering Statistics
IMSE 4410 Spring 2015. 
Copyright 2013-2015 by Timothy Middelkoop License CC by SA 3.0

## Multiple Linear Regression (Chapter 12, Week 11)

```{r}
ch12 <- read.csv("data/5e/ch12.csv",header=TRUE)
nrow(ch12)
```

### Example 12.1 Wire Bond Strength.

```{r}
## note the c column and the order used later.
d <- na.omit(data.frame(y=ch12$Example12.1S,c=1,x1=ch12$Example12.1L,x2=ch12$Example12.1H))
n <- nrow(d) ; n

plot(d)

## Model
m <- lm(y~x1+x2,d) ; s <- summary(m) ; s
plot(m)

## y:(n,1) X:(n,p) betah:(p,1) e:(n,1)
x <- data.matrix(d[2:4]) ; x[0:5,]

y <- data.matrix(d[1]) ; y[0:5]

bh = solve(t(x) %*% x) %*% t(x) %*% y ; bh

## yh and e, verify with text.
yh <- x %*% bh ; yh[0:5,]

e <- y-yh ; e[0:5,]

## three parameters (p=k+1)
p <- ncol(d)-1 ; p # [1] 3

## Estimating σ² (sigma^2), sigmahsq
sse <- sum(e^2) ; sse # [1] 115.1735

sigmahsq <- sse/(n-p) ; sigmahsq # [1] 5.235158

s$sigma^2 # [1] 5.235158
```

### Section 12-2 Hypothesis Tests in Multiple Linear Regression

```{r}
## R² (multiple R-sqared)
yb <- mean(d$y) ; yb # [1] 29.0328

ssr <- sum((yh-yb)^2) ; ssr # [1] 5990.771

sst <- sum((y-yb)^2) ; sst # [1] 6105.945
sse+ssr # [1] 6105.945

rsq <- ssr/sst ; rsq # [1] 0.9811375

s$r.sq # [1] 0.9811375

## Check with anova
a <- anova(m) ; a

names(a)

sum(a$Sum) # [1] 6105.945

## Mean Squared corrects for the DOF
mse <- sse/(n-p) ; mse # [1] 5.235158

## Tests on individual parameters
#s
# Residual standard error: 2.288 on 22 degrees of freedom
# Multiple R-squared:  0.9811,	Adjusted R-squared:  0.9794 
# F-statistic: 572.2 on 2 and 22 DF,  p-value: < 2.2e-16

## See text for test statistic, same as last time.
## t-value, H0:betai=0, H1:betai!=0

## Significance of regression (p=k+1)
## H0:b1=b2=b3...=0 H1:b1!=0 for at least one.
msr <- ssr/(p-1) ; msr
mse <- sse/(n-p) ; mse
f0 <- msr/mse ; f0 # [1] 572.1672
```

* What are the F-stats in the ANOVA though?
* It is the test of the significance of adding a variable
* it is equivalent to the t-test.

### 12-3 Confidence Intervals on multiple parameters.

* In principal how is this found? Specifically where are the equations?
* How is this used?
* Undergraduate: read and follow this section in the text.
* Graduate/Extra: compute and verify the results below using the text.

```{r}
## Confidence Intervals on the regression coefficients.
confint(m)

## Confidence Interval on the mean response
predict(m, data.frame(x1=8,x2=275))

```
